---
title: "Proposal title"
subtitle: "Proposal"
author: 
  - name: "Datavista - Team member 1, Team member 2,..."
    affiliations:
      - name: "School of Information, University of Arizona"
description: "Project description"
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
---

# High Level Goal

The goal of the project is to visualise

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false
```

A brief description of your dataset including its provenance, dimensions, etc. as well as the reason why you chose this dataset.

Make sure to load the data and use inline code for some of this information.

## Questions

The two questions you want to answer.

## Analysis plan

# Plan of Attack

This document outlines the tasks for our project, including their status, assignees, due dates, priorities, and summaries.

| Task Name                                               | Status                            | Assignee | Due        | Priority | Summary                                     |
|------------|------------|------------|------------|------------|------------|
| Dataset and question selection                          | Explored and finalize the dataset | Everyone | March 24th | HIGH     | Selected dataset and question to be solved. |
| Brainstorming and working on data analysis and proposal | Completed the proposal            | Everyone | April 1st  | High     | Proposal ready for review.                  |
|                                                         |                                   | Everyone | April 8th  | High     |                                             |
|                                                         |                                   |          | April 15th |          |                                             |
|                                                         |                                   |          | April 22nd |          |                                             |
|                                                         |                                   |          | April 29th |          |                                             |
|                                                         |                                   |          | May 6th    |          |                                             |

```         
                                       |
```

# Repo Organization

The following folders comprise the project repository

-   **.github/:** This directory is designated for files associated with GitHub, encompassing workflows, actions, and templates tailored for issues.

-   **\_extra/:** Reserved for miscellaneous files that don't neatly fit into other project categories, providing a catch-all space for various supplementary documents.

-   **\_freeze/:** Within this directory lie frozen environment files containing comprehensive information regarding the project's environment configuration and dependencies.

-   **data/:** Specifically allocated for storing indispensable data files crucial for the project's functionality, encompassing input files, datasets, and other essential data resources.

-   **images/:** Serving as a repository for visual assets employed throughout the project, including diagrams, charts, and screenshots, this directory maintains visual elements integral to project documentation and presentation.

-   **.gitignore:** This file functions to specify exclusions from version control, ensuring that designated files and directories remain untracked by Git, thus streamlining the versioning process.

-   **README.md:** Serving as the primary hub of project information, this README document furnishes essential details encompassing project setup, usage instructions, and an overarching overview of project objectives and scope.

-   **\_quarto.yml:** Acting as a pivotal configuration file for Quarto, this document encapsulates various settings and options governing the construction and rendering of Quarto documents, facilitating customization and control over document output.

-   **about.qmd:** This Quarto Markdown file supplements project documentation by providing additional contextual information, elucidating project purpose, contributor insights, and other pertinent project details.

-   **index.qmd:** This serves as the main documentation page for our project. This Quarto Markdown file provides detailed descriptions of our project, including code examples and visualizations.

```{}
```
